---
title: "scicomptools Vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{scicomptools Vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r knitr-mechanics, include = F}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r pre-setup, echo = F, message = F}
# install.packages("devtools")
# devtools::install_github("NCEAS/scicomptools")
```

## Overview

The `scicomptools` package is built to house self-contained tools built by the NCEAS Scientific Computing Team. These functions are written based on the needs of various synthesis working groups so each function operates more or less independently of the others. We hope these tools will be valuable to users in and outside of the context for which they were designed!

This vignette describes some of the main functions of `scicomptools` using the examples.

```{r setup}
library(scicomptools)
```

### Extract Summary Statistics Tables



### Create Google Drive Table of Contents



### Read in Multiple Excel Sheets

Working groups often store their data in Microsoft Excel files with multiple sheets (often one sheet per data type or per collection location). We wrote `read_xl_sheets` to import every sheet in a user-specified MS Excel file and store each sheet in a list. If every sheet has the same columns, you can then easily unlist them into a flat dataframe (rather than the somewhat laborious process of reading in each sheet separately).

`read_xl_sheets` only has a single argument (`file_name`) which accepts the name of (and path to) the MS Excel file to read.

```{r read_xl_sheets}
# Read in sheets
sheet_list <- scicomptools::read_xl_sheets(file_name = system.file("extdata", "faux_data.xlsx", 
                                                     package = "scicomptools"))

# Show structure
dplyr::glimpse(sheet_list)
```

We also wrote a companion function named `read_xl_format` where the specific formatting of all *cells* in the sheets of an Excel file is extracted. This is useful if fill color or text formatting is used to denote information that is not redundant with columns (i.e., information that is lost when reading an Excel sheet into an API). `read_xl_format` uses the same syntax as `read_xl_sheets` to maximize interoperability.

```{r read_xl_format}
# Read in *format of* sheets
form_list <- scicomptools::read_xl_format(file_name = system.file("extdata", "faux_data.xlsx", 
                                                     package = "scicomptools"))

# Show structure of that
dplyr::glimpse(form_list)
```

### Handling Two Working Directories

Many of the working groups that we support work both on local computers as well as in remote servers. We leverage Git and GitHub to pass code updates back and forth between the two environments but updating the working directory every time such a pivot is made is cumbersome.

We have thus written `wd_loc` for this purpose! `wd_loc` allows users to specify two working directory paths and toggle between them using a logical argument.

The `local` argument accepts a logical value. If `local = TRUE`, then the file path specified in the `local_path` argument is returned as a character string. If `local = FALSE` then the file path specified in the `remote_path` argument is used. `local_path` defaults to `base::getwd()` so it need not be specified if you are using RStudio Projects.

You can run `wd_loc` inside of `base::setwd` if desired though we recommend assigning the file path to a "path" object and invoking that whenever import or export must be done.

```{r wd_loc, eval = F}
scicomptools::wd_loc(local = TRUE,
                     local_path = getwd(),
                     remote_path = file.path("path on server"))
```

```{r wd_loc message, echo = F}
message("/Users/.../scicomptools/vignettes")
```

### Checking Tokens

Some operations require passing a user's personal access token (a.k.a. "PAT") to RStudio. These workflows can fail unexpectedly if the token is improperly set or expires so it is valuable to double check whether the token is still embedded.

The `token_check` function checks whether a token is attached and messages whether or not one is found. Currently this function supports checks for Qualtrics and GitHub tokens but this can be expanded if need be ([post a GitHub issue](https://github.com/NCEAS/scicomptools/issues) if you have another token in mind!). Additionally, if `secret = TRUE` (the default) the success message simply identifies whether a token is found. If `secret = FALSE` the success message prints the token string in the console.

```{r token_check}
scicomptools::token_check(api = "github", secret = TRUE)
```
